Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated

### 2024-10-22 submission/new_manuscript/manuscript.tex
changes made to new_manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{attractiveness, AI-generated images, fiction, fake news, sense of reality\newline\indent Word count: 5242}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated},
  pdfauthor={Dominique Makowski1, An Shu Te2, Stephanie Kirk2, Ngoi Zi Liang2, Panagiotis Mavros3, \& S.H. Annabel Chen2, 4, 5, 6},
  pdflang={en-EN},
  pdfkeywords={attractiveness, AI-generated images, fiction, fake news, sense of reality},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}}
\author{Dominique Makowski\textsuperscript{1}, An Shu Te\textsuperscript{2}, Stephanie Kirk\textsuperscript{2}, Ngoi Zi Liang\textsuperscript{2}, Panagiotis Mavros\textsuperscript{3}, \& S.H. Annabel Chen\textsuperscript{2, 4, 5, 6}}
\date{}

\shorttitle{Attractiveness and Reality}

\authornote{

Correspondence concerning this article should be addressed to Dominique Makowski (\href{mailto:D.Makowski@sussex.ac.uk}{\nolinkurl{D.Makowski@sussex.ac.uk}}).

The authors made the following contributions. Dominique Makowski: Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Software, Supervision, Validation, Visualization, Writing -- original draft; An Shu Te: Data curation, Project administration, Resources, Investigation, Writing -- original draft; Stephanie Kirk: Project administration, Resources, Writing -- original draft; Ngoi Zi Liang: Project administration, Resources, Writing -- review \& editing; Panagiotis Mavros: Supervision, Writing -- review \& editing; S.H. Annabel Chen: Project administration, Supervision, Writing -- review \& editing.

Correspondence concerning this article should be addressed to Dominique Makowski, School of Psychology, University of Sussex, Brighton. E-mail: \href{mailto:D.Makowski@sussex.ac.uk}{\nolinkurl{D.Makowski@sussex.ac.uk}}

@@

\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Psychology, University of Sussex, UK\\\textsuperscript{2} School of Social Sciences, Nanyang Technological University, Singapore\\\textsuperscript{3} Department of Economics and Social Sciences, Télécom Paris, France\\\textsuperscript{4} LKC Medicine, Nanyang Technological University, Singapore\\\textsuperscript{5} National Institute of Education, Singapore\\\textsuperscript{6} Centre for Research and Development in Learning, Nanyang Technological University, Singapore}

\abstract{%
Technological advances render the distinction between artificial (e.g., computer-generated faces) and real stimuli increasingly difficult, yet the factors driving our beliefs regarding the nature of ambiguous stimuli remain largely unknown. In this study, 150 participants rated 109 pictures of faces on 4 characteristics (attractiveness, beauty, trustworthiness, familiarity). The stimuli were then presented again with the new information that some of them were AI-generated, and participants had to rate each image according to whether they believed them to be real or fake. Despite all images being pictures of real faces from the same database, most participants did indeed rate a large portion of them as ``fake'' (often with high confidence), with strong intra- and inter-individual variability. Our results suggest a gender-dependent role of attractiveness on reality judgements, with faces rated as more attractive being classified as more real. We also report links between reality beliefs tendencies and dispositional traits such as narcissism and paranoid ideation.
}

\begin{document}
\maketitle

For the first time in human history, technology has enabled the creation of near-perfect simulations indistinguishable from reality. These artificial, yet realistic constructs permeate all areas of life through immersive works of fiction, deep fakes (real-like images and videos generated by deep learning algorithms), virtual and augmented reality (VR and AR), artificial beings (artificial intelligence ``bots'' with or without a physical form), fake news and skewed narratives, of which ground truth is often hard to access (Nightingale \& Farid, 2022). Such developments not only carry important consequences for the technological and entertainment sectors, but also for security and politics - for instance if used for propaganda and disinformation, recruitment into malevolent organizations, or religious indoctrination (Pantserev, 2020). This issue is central to what has been coined the ``post-truth era'' (Lewandowsky et al., 2017), in which the distinction (and lack thereof) between authentic and simulated objects will play a critical role.

While not all simulations have achieved perfect realism (Corvi et al., 2022; e.g., Computer Generated Images - CGI in movies or via recent algorithms such as GANs or diffusion models often include distortions or lack certain key details that makes them visually distinct from real images, McDonnell \& Breidt, 2010), it is fair to assume that these technical limitations will become negligible in the near future. This is particularly true in the field of face generation, where face-generation algorithms are already able to create stimuli that are virtually indistinguishable from real photos (Moshel et al., 2022; Nightingale \& Farid, 2022; Tucciarelli et al., 2020). Such a technological feat, however, leads to a new question: if real and fake stimuli cannot be differentiated based on their objective ``physical'' characteristics, how can we form judgements regarding their nature?

Literature shows that the context surrounding a stimulus often plays an important role in the assessment of its reality (a process henceforth referred to as \emph{simulation monitoring}, Makowski, 2018; Makowski, Sperduti, et al., 2019). With the extensive search and processing of cues within ambiguous stimuli being an increasingly complex and cognitively effortful strategy (Michael \& Sanson, 2021; Susmann et al., 2021), people tend to draw on peripheral contextual cues (\textbf{Figure 1}), such as the source of the stimulus (e.g., which journal was the information published in), and its credibility, authority and expertise, to help facilitate their evaluation (Michael \& Sanson, 2021; Petty \& Cacioppo, 1986; Susmann et al., 2021). However, the atomization and decontextualization of information allowed by online social media (where text snippets or video excerpts are often mass-shared with little context) makes this task progressively difficult (Berghel, 2018; Y. Chen et al., 2015). Thus, in the absence of clear contextual information, what drives our beliefs of reality?

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/Figure1} \caption{The decision to believe that an ambiguous stimulus (of any form, e.g., images, text, videos, environments, ...) is real or fake depends of individual characteristics (e.g., personality and cognitive styles), stimulus-related features (context, emotionality), and their interaction, which can manifest for instance in our bodily reaction.}\label{fig:unnamed-chunk-2}
\end{figure}

Evidence suggests that inter-individual characteristics play a crucial role in simulation monitoring, with factors such as cognitive style, prior beliefs, and personality traits (Bryanov \& Vziatysheva, 2021; Ecker et al., 2022; Sindermann et al., 2020). For instance, individuals with stronger analytical reasoning skills have been found to better discriminate real from fake stimuli (Pehlivanoglu et al., 2021; Pennycook \& Rand, 2019), and prior knowledge or beliefs about the stimulus influences one's perception of it by biasing the attention deployment towards information that is in line with one's expectations (Britt et al., 2019). Furthermore, dispositional traits, such as high levels of narcissism and low levels of openness and conscientiousness, have been associated with greater susceptibility to fake news (Piksa et al., 2022; Sindermann et al., 2020).

Beyond stimulus- and individual-related characteristics, evidence suggests that the interaction between the two (i.e., the subjective reaction associated with the experience of a given stimulus), contributes to simulation monitoring decisions. For instance, the intensity of experienced emotions have been shown to increase one's sense of presence - the extent to which one feels like ``being there'', as if the object of experience was real - when engaged in a fictional movie or a VR environment (Makowski et al., 2017; Sanchez-Vives \& Slater, 2005). Indeed, participants' self-reported emotional arousal were found to significantly predict the probability that they would perceive images as real (Azevedo et al., 2020). Conversely, beliefs that emotional stimuli were fake (e.g., that emotional scenes were not authentic but instead involved actors and movie makeup) were found to result in emotion down-regulation (Makowski, Sperduti, et al., 2019; Sperduti et al., 2017). In line with these findings, studies on susceptibility to fake news have also found heightened stimulus emotionality to be associated with greater belief (Bago et al., 2022; Martel et al., 2020), and higher neurophysiological arousal was predictive of judging realistic images as real (Azevedo et al., 2020). Additionally, other factors, such as the stimuli's perceived self-relevance (Goldstein, 2009; Sperduti et al., 2016), as well as familiarity (Begg et al., 1992), could also play a role in guiding our appraisal of a stimulus.

>>>>

\documentclass[
  man,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[bidi=default]{babel}
\babelprovide[main,import]{english}

% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}

\RequirePackage{longtable}
% \setlength\LTleft{0pt}
\RequirePackage{threeparttablex}

% %

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[nolongtablepatch]{lineno}
\linenumbers

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{Too Beautiful to be Fake: Attractive Faces are Less Likely to be
Judged as Artificially Generated}
\shorttitle{Attractiveness and Reality}

\usepackage{etoolbox}

\authorsnames[{1},{2},{2},{2},{3},{1,4,5,6}]{Dominique Makowski,An Shu
Te,Stephanie Kirk,Ngoi Zi Liang,Panagiotis Mavros,S.H. Annabel Chen}

\authorsaffiliations{
{School of Psychology, University of Sussex},{School of Social
Sciences, Nanyang Technological University},{Department of Economics and
Social Sciences, Télécom Paris},{LKC Medicine, Nanyang Technological
University},{National Institute of Education},{Centre for Research and
Development in Learning, Nanyang Technological University}}

\leftheader{Makowski, Te, Kirk, Liang, Mavros and Chen}

\abstract{Technological advances render the distinction between
artificial (e.g., computer-generated faces) and real stimuli
increasingly difficult, yet the factors driving our beliefs regarding
the nature of ambiguous stimuli remain largely unknown. In this study,
150 participants rated 109 pictures of faces on 4 characteristics
(attractiveness, beauty, trustworthiness, familiarity). The stimuli were
then presented again with the new information that some of them were
AI-generated, and participants had to rate each image according to
whether they believed them to be real or fake. Despite all images being
pictures of real faces from the same database, most participants did
indeed rate a large portion of them as `fake' (often with high
confidence), with strong intra- and inter-individual variability. Our
results suggest a gender-dependent role of attractiveness on reality
judgements, with faces rated as more attractive being classified as more
real. We also report links between reality beliefs tendencies and
dispositional traits such as narcissism and paranoid ideation.}
% 
\keywords{attractiveness, AI-generated images, fiction, fake news, sense
of reality}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{An Shu
Te}{0000-0002-9312-5552}}\par{\addORCIDlink{Stephanie
Kirk}{0000-0002-9312-5552}}\par{\addORCIDlink{Panagiotis
Mavros}{0000-0002-1540-5516}}
\par{ }
\par{       Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Dominique
Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft; An Shu Te:   Project
administration, Resources, Writing -- original draft; Stephanie
Kirk:   Project administration, Resources, Writing -- original
draft; Ngoi Zi Liang:   Project administration, Resources, Writing --
review \& editing; Panagiotis Mavros:   Supervision, Writing -- review
\& editing; S.H. Annabel Chen:   Project
administration, Supervision, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique Makowski, Email: D.Makowski@sussex.ac.uk}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\resetlinenumber[1]

Advancements in technology have now made it possible to create
near-perfect simulations that are indistinguishable from reality with an
ease, affordability and accessibility that are unprecedented in Human
history.These artificial, yet realistic constructs permeate all areas of
life through immersive works of fiction, deep fakes (real-like images
and videos generated by deep learning algorithms), virtual and augmented
reality (VR and AR), artificial beings (artificial intelligence ``bots''
with or without a physical form), fake news and skewed narratives, of
which ground truth is often hard to access (Nightingale \& Farid, 2022).
Such developments not only carry important consequences for the
technological and entertainment sectors, but also for security and
politics - for instance if used for propaganda and disinformation,
recruitment into malevolent organizations, or religious indoctrination
(Pantserev, 2020). This issue is central to what has been coined the
``post-truth era'' (Lewandowsky et al., 2017), in which the distinction
(and lack thereof) between authentic and simulated objects will play a
critical role.

While not all simulations have achieved perfect realism, such as
Computer Generated Images (CGI) in movies or via recent algorithms such
as GANs or diffusion model, which often include distortions or lack
certain key details distinguishing them from real images (Corvi et al.,
2022; McDonnell \& Breidt, 2010), it is fair to assume that these
technical limitations will become negligible in the near future. This is
particularly true in the field of face generation, where face-generation
algorithms are already able to create stimuli that are virtually
indistinguishable from real photos (Moshel et al., 2022; Nightingale \&
Farid, 2022; Tucciarelli et al., 2020). Such a technological feat,
however, leads to a new question: if real and fake stimuli cannot be
differentiated based on their objective ``physical'' characteristics,
how can we form judgements regarding their nature?

Literature shows that the context surrounding a stimulfus often plays an
important role in the assessment of its reality (a process henceforth
referred to as \emph{simulation monitoring}, Makowski, 2018; Makowski,
Sperduti, et al., 2019). With the extensive search and processing of
cues within ambiguous stimuli being an increasingly complex and
cognitively effortful strategy (Michael \& Sanson, 2021; Susmann et al.,
2021), people tend to draw on peripheral contextual cues (\textbf{Figure
1}), such as the source of the stimulus (e.g., which journal was the
information published in), and its credibility, authority and expertise,
to help facilitate their evaluation (Michael \& Sanson, 2021; Petty \&
Cacioppo, 1986; Susmann et al., 2021). However, the atomization and
decontextualization of information allowed by online social media (where
text snippets or video excerpts are often mass-shared with little
context) makes this task progressively difficult (Berghel, 2018; Y. Chen
et al., 2015). Thus, in the absence of clear contextual information,
what drives our beliefs of reality?

\begin{figure}[H]

\caption{Figure 1. The decision to believe that an ambiguous stimulus
(of any form, e.g., images, text, videos, environments, \ldots) is real
or fake depends of individual characteristics (e.g., personality and
cognitive styles), stimulus-related features (context, emotionality),
and their interaction, which can manifest for instance in our bodily
reaction.}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../figures/figure1.png}

@@

\end{figure}%

Evidence suggests that inter-individual characteristics play a crucial
role in simulation monitoring, with factors such as cognitive style,
prior beliefs, and personality traits (Bryanov \& Vziatysheva, 2021;
Ecker et al., 2022; Sindermann et al., 2020). For instance, individuals
with stronger analytical reasoning skills have been found to better
discriminate real from fake stimuli (Pehlivanoglu et al., 2021;
Pennycook \& Rand, 2019), and prior knowledge or beliefs about the
stimulus influences one's perception of it by biasing the attention
deployment towards information that is in line with one's expectations
(Britt et al., 2019). Furthermore, dispositional traits, such as high
levels of narcissism and low levels of openness and conscientiousness,
have been associated with greater susceptibility to fake news (Piksa et
al., 2022; Sindermann et al., 2020). \textbf{Interestingly, a recent
review suggested that narcissism was related to a strong self-perceived
ability at detecting lies (Turi et al., 2022), which could translate to
participants scoring high on narcissism providing more clear cut and
confidence responses. Conversely, those high in honesty-humility tend to
be more conservative in their judgments to ensure fairness (Liu et al.,
2020), likely resulting in lower confidence ratings.}

Beyond stimulus- and individual-related characteristics, evidence
suggests that the interaction between the two (i.e., the subjective
reaction associated with the experience of a given stimulus),
contributes to simulation monitoring decisions. For instance, the
intensity of experienced emotions have been shown to increase one's
sense of presence - the extent to which one feels like ``being there'',
as if the object of experience was real - when engaged in a fictional
movie or a VR environment (Makowski et al., 2017; Sanchez-Vives \&
Slater, 2005). Indeed, participants' self-reported emotional arousal
were found to significantly predict the probability that they would
perceive images as real (Azevedo et al., 2020). Conversely, beliefs that
emotional stimuli were fake (e.g., that emotional scenes were not
authentic but instead involved actors and movie makeup) were found to
result in emotion down-regulation (Makowski, Sperduti, et al., 2019;
Sperduti et al., 2017). In line with these findings, studies on
susceptibility to fake news have also found heightened stimulus
emotionality to be associated with greater belief (Bago et al., 2022;
Martel et al., 2020), and higher neurophysiological arousal was
predictive of judging realistic images as real (Azevedo et al., 2020).
Additionally, other factors, such as the stimuli's perceived
self-relevance (Goldstein, 2009; Sperduti et al., 2016), as well as
familiarity (Begg et al., 1992), could also play a role in guiding our
appraisal of a stimulus.\textbf{For instance,} Miller et al. (2023)
\textbf{reported that participants were more likely to mistakenly
identify AI-generated faces as real because they perceived them as more
familiar.}

Due to their popularity as a target of CGI technology and the prospect
offered with facial features that can be experimentally manipulated,
AI-generated images of faces are increasingly used to study face
processing (Dawel et al., 2021), in particular in relationship with
saliency or emotions, as well as to other important components of face
evaluation, such as trustworthiness or attractiveness (Balas \& Pacella,
2017; Calbi et al., 2017; Sobieraj \& Krämer, 2014; Tsikandilakis et
al., 2019). Interestingly, artificially created faces rated as more
attractive (by an independent group of raters) were perceived as less
real (Tucciarelli et al., 2020). Conversely, Liefooghe et al. (2022)
reports that attractiveness ratings were significantly lower when
participants who were told that the faces were AI-generated were
compared to those who had no prior knowledge. \textbf{Similarly, when
participants are informed that faces are AI-generated, the perceived
artificiality leads to lower trust ratings (Wang \& Nishida, 2024), even
when they are real faces (Liefooghe et al., 2022). In contrast, when
participants are unaware that the faces are AI-generated, trust ratings
for these synthetic faces tend to increase (Nightingale \& Farid,
2022)}. Whereas this line of evidence suggests that reality beliefs have
an effect on face attractiveness and \textbf{trustworthiness} ratings,
the opposite question - whether attractiveness \textbf{and
trustworthiness contribute} to the formation of reality beliefs - has
received little attention to date.

AI-generated content, in particular realistic images, is becoming
commonplace and carries important risks for misinformation and
black-mailing (Viola \& Voto, 2023), emphasizing the need to understand
the different components that come into play in the formation of reality
beliefs. This exploratory study primarily aims at investigating the
effect of facial attractiveness on simulation monitoring, i.e., on the
beliefs that an image is real or artificially generated. Based on the
\emph{affective reality theory} (Makowski, 2018, 2023), which suggests
that salient and emotional stimuli are perceived to be more real (up to
a point of reversal after which beliefs of fiction becomes used an
emotion regulation strategy), we hypothesize a quadratic relationship
between perceived realness and attractiveness: faces rated as highly
attractive or unattractive will more likely be believed to be real. We
expect a similar relationship with trustworthiness ratings given its
well-established link with attractiveness (Bartosik et al., 2021;
Garrido \& Prada, 2017; Liefooghe et al., 2022; Little et al., 2011),
and a positive relationship with familiarity (as more familiar faces
would appear as more salient, self-relevant and anchored in reality).
Additionally, we will further explore the link shared by dispositional
traits, such as personality and attitude towards AI, with simulation
monitoring tendencies. Importantly, this study does investigate the
discriminative accuracy between ``true'' photos and ``true''
artificially-generated images (which we consider more a technological
issue than a psychological one), focusing on the beliefs that a stimulus
is real or fake, independently of its true. \textbf{In other words, the
present study investigates the psychological process that leads to
different \emph{beliefs} of reality, rather than the discrimination
between real faces and actual AI-generated ones, which largely depends
on the technological quality of the AI-generation process.}

\subsection{Methods}\label{methods}

All the material (preregistration \textbf{\emph{{[}FOOTNOTE: This
approach diverges from the preregistration in several key ways. First,
the phrasing of items was modified from ``Assuming the face you saw was
of a real individual, how\ldots{}'' to ``I find this person\ldots{}'' A
new attractiveness scale (i.e., Beauty) was introduced to capture a more
objective measure of attraction. Finally, the data analysis method was
altered from Bayesian Mixed Models due to computational limitations, as
we were unable to run these models on a high-performance cluster.{]}}},
experiment demo, experiment code, raw data, analysis script with
complementary figures and analyses, etc.) is available at
\href{https://github.com/RealityBending/FakeFace}{\textbf{https://github.com/RealityBending/FakeFace}}.

\textbf{Ethics Statement.} This study was approved by the NTU
Institutional Review Board (NTU IRB-2022-187) and all procedures
performed were in accordance with the ethical standards of the
institutional board and with the 1964 Helsinki Declaration. All
participants provided their informed consent prior to participation and
were incentivized after completing the study.
